### 2009-12-11 docs/marcello-dissertation/thesis/chapter4-netbeams-in-depth-requirements-analysis-and-specification.tex
Reviews of chapter 4 of the report... Adding a label on chapter 2 and adding references to non-cited sections.
git-svn-id: svn://localhost/netbeams/branches/marcello/persistence@689 559edde2-a4d4-44e0-96fc-4035e40d7396
Review of the audio feedback from Prof. Puder.
- Niki suggested the use of in, as Prof. Puder asked to verify;
- Adding the name of the YSI Sonde, referring to the manufacturer YSI Incorporated
- Adding the dimensions information of the device, as well as the name some properties collected
- Fixing the table reference, actually changing the data stream from table to Listing
- Ad

As discussed in Chapter 2, different properties of a sensor network and the
nature of the collected data must be taken into account in order to provide a
data persistence layer for a given sensor network. Similarly, in order to
better assist one's analysis of such functionality, Chapter 3 proposed a set of
data persistence taxonomies related to different properties of the collected
data life cycle. The goal of this chapter is to describe the fundamental properties 
of the case study used for this work, that is, NetBEAMS, in order to lay the 
groundwork as it relates to the selection of the database technology. NetBEAMS 
provides an automated infrastructure solution for SF-BEAMS, and this last component 
will be covered first. Then, this work proposes a categorization of NetBEAMS to 
the face of the taxonomies.

@@

As described by \cite{netbeams2009}, NetBEAMS is a joint venture between the
department of Biology and Computer Science at San Francisco State University,
whose goal is to automate the operational execution of SF-BEAMS. The San Francisco
Bay Environmental Assessment and Monitoring Station, or SFBEAMS \cite{sfbeams2006}, 
is an environmental sensor network system whose primary focus is the study of complex 
marine and estuarine environments using the SF-BEAMS sensors. The sensors are deployed off a
pier located on the San Francisco Bay, in Tiburon, California; and is operated by 
the Romberg Tiburon Center for Environmental Studies (RTC) a part of San Francisco
State University. NetBEAMS is an experimental component that utilizes the SF-BEAMS 
infrastructure to operate. This section details SF-BEAMS in general, and presents 
the requirements for data persistence.

@@

In general, the SF-BEAMS network infrastructure contains a varying number of
wired and wireless devices attached to pylons at the pier. Each of them
is responsible for observing different conditions of the area, having its own
mechanisms for internal storage for the collected data. In this way, data can
be directly transferred to the labs via Ethernet cables or collected manually
with a laptop computer. The current data collection process for SF-BEAMS is
described in Figure
\ref{fig:SF-BEAMS-system-architecture}\cite{sfbeams-current-system}. Upon
collecting data from sensors, the RTC staff uses automation scripts written in
Matlab \cite{matlab} to process, index and distribute the raw data in
different formats. One of these formats is the OPeNDAP \cite{opendap}, which
is widely used at research institutions to promote easier data exchange
amongst them. The SF-BEAMS website provides access to the collected data through
the Internet at http://sfbeams.sfsu.edu:8080/opendap. An example of an HTTP
Request from accessing the data for a specific YSI \cite{YSI-Sonde} is shown in
Listing \ref{file:rtc-ysi-opendap}.

@@

As described in section \ref{sec:sn-infrastructure}, sensor devices produce
the observed data based on properties of measurements defined by its
manufacturer. NetBEAMS has used ``YSI 6600 ESD V2" \cite{YSI-Sonde}, as seen in
Figure \ref{fig:ysi-device}, one of the sensors devices used by SF-BEAMS. It is a
powerful water quality-monitoring device, capable of producing around 52 bytes
on a single real-time data stream reading, as shown on Table
\ref{tab:ysi-data-stream}.

@@

\begin{table}
    \begin{center}
        \begin{tabular}{|l|}\hline
  12.20    192    179 5588.40   0.09   0.084   0.059  7.98   -79.6   99.5  8.83  0.4     8.7\\\hline
        \end{tabular}
    \end{center}
    \label{tab:ysi-data-stream}
    \caption{The Collected Data Stream from the YSI Sonde 6600ESDV2}
\end{table}

In order to have access to the observed data, the network staff downloads the
data using one of the device's connections such as the RS-232 serial connector
\cite{rs232}. Then, they upload the data into their data servers where they are
processed and archived for historical analysis. According to one
staff member, the infrastructure of the SF-BEAMS is comprised of the 
following components (as of May 2009):

@@

Considering the infrastructure, the data load in the server-side
of RTC's laboratories produced by the YSI Sonde sensors devices can be
estimated by analyzing the data distribution over a period of one year, as
shown in Table \ref{tab:ysi-data-distribution}:

\begin{table}[!b]
    \label{tab:ysi-data-distribution}
     \begin{center}
      \begin{tabular}{|c|c|c|c|c|c|c|}\hline 
        \textbf{YSIs} & \textbf{Rate} & \textbf{Hourly} & \textbf{Daily} &
        \textbf{Weekly} & \textbf{Monthly} & \textbf{Yearly}\\\hline 
        1 & 1 min & 3.04 Kb & 73.12 Kb & 511.87 Kb & 1.99 Mb & 23.99 Mb\\\hline 
        5 & 6 min & 15.23 Kb & 365.62 Kb & 2.5 Mb & 9.99 Mb & 119.97 Mb\\\hline 
        1 & 15 min & 0.5 Kb & 12.18 Kb & 85.31 Kb & 341.25 Kb & 3.99 Mb\\\hline 
        5 & 1 min & 2.54 Kb & 60.93 Kb & 426.56 Kb & 1.67 Mb & 19.99 Mb\\\hline
        1 & 6 min & 0.2 Kb & 4.87 Kb & 34.12 Kb & 136.5 Kb & 1.6 Mb\\\hline 
        5 & 15 min & 1.0 Kb & 24.37 Kb & 170.62 Kb & 682.5 Kb & 7.99 Mb\\\hline
        \end{tabular}
      \end{center}
    \caption{Amount of data produced by the RTC's YSI sondes}
\end{table}

@@

\begin{itemize}
  \item \textbf{The Purpose of Sensor Data}: the data from SF-BEAMS is exclusively
  used for the purpose of Data Archival, since they are stored in RTC's lab;
  \item \textbf{The Location of the Sensor Data}: it is clear that the purpose
  of the data is directly related to the sensors location and, thus, the
  strategy of \textbf{External Data Storage} is used to store its collected
  data at the network sink at the RTC site;
  \item \textbf{Data Model}: since the format used by the RTC staff to share
  data with researchers is in the OPeNDAP format, the data
  model used is the tabular data model, since it uses the comma-delimited
  files. In this way, the \textbf{Schema-less} model is used;
  \item \textbf{Data Description}: once the collected data reaches the RTC
  lab, a quality assurance process is executed, and the data is transformed
  into the format used by OPeNDAP. This conversion carries data regarding
  the time of the data collection, therefore, one of the
  \textbf{Time Dimensions} such as the valid time is used. Since the files are stored
  with specific file name structure, using timestamps to describe the data, the
  use of transaction time may be also considered. Finally, the contents of the
  files contain the \textbf{Data Identity} as shown in the samples of
  the collected in Listing \ref{file:rtc-ysi-opendap};
  \item \textbf{Query Processing Mechanism}: similar to the data stored in a
  centralized way, the query processing is also \textbf{Centralized}, as the
  data is stored on file system;
  \item \textbf{Data Volume}: the volume of data produced is characterized by small.
 It is related to the volume of data produced by the sensor devices,
  which is numerical data streams, and the total number of sensor devices;
  \item \textbf{System Organization}: SF-BEAMS runs on a Single System, as
  the collected data is stored directly in the file-system after previous 
  processing. However, the collected data is provided to the Internet using
  the OPeNDAP Hyrax system as a middleware.
\end{itemize}

@@

Although the execution of RTC's sensor network can be operated as it
was mentioned in the previous sections, \cite{netbeams2009} described some of the
operational challenges faced by the RTC staff during regular activities of the
data collection process. It was clear to the research group that SF-BEAMS
could have not only its data gathering process automated, but also its data
management and distribution. By the use of COTS\footnote{Common-Off-The-Shelf
designates a product that is produced and sold in bulk} embedded devices and
open-source \cite{open-source} software, the research group developed a second
version of NetBEAMS, a component-based approach which suggested the improvement of 
operational activities from SF-BEAMS using systems automation. In brief, one of the 
ongoing problems of NetBEAMS was regarding Data Persistence, the primary motivation of
this report.

@@

The scope of this work is defined as to provide a persistence layer for
NetBEAMS, making a selection of a persistence layer that is an exception to the
taxonomical classification of SF-BEAMS. Furthermore, the technology to be
selected for the persistence layer must take into account the requirements of
NetBEAMS primary users such as Computer Science researchers, as well as end
users such as Biologists from the RTC laboratories. The requirements for the
data persistence specifically for NetBEAMS are described in the following
section.

>>>>

As discussed in Chapter 2, different properties of a sensor network and the
nature of the collected data must be taken into account in order to provide a
data persistence layer for a given sensor network. Similarly, in order to
better assist one's analysis of such functionality, Chapter 3 proposed a set of
data persistence taxonomies related to different properties of the collected
data life cycle. The goal of this chapter is to describe the fundamental properties 
of the case study used for this work, that is, NetBEAMS, in order to lay the 
groundwork as it relates to the selection of the database technology. NetBEAMS 
provides an automated infrastructure solution for SF-BEAMS, and this last component 
will be covered first. Then, this work proposes a categorization of NetBEAMS
according to the defined taxonomies.

@@

As described by \cite{netbeams2009}, NetBEAMS is a joint venture between the
department of Biology and Computer Science at San Francisco State University,
whose goal is to automate the operational execution of SF-BEAMS. The San Francisco
Bay Environmental Assessment and Monitoring Station, or SFBEAMS \cite{sfbeams2006}, 
is an environmental sensor network system whose primary focus is the study of complex 
marine and estuarine environments using the SF-BEAMS sensors. The sensors are deployed off a
pier located in the San Francisco Bay, in Tiburon, California; and is operated by 
the Romberg Tiburon Center for Environmental Studies (RTC) a part of San Francisco
State University. NetBEAMS is an experimental component that utilizes the SF-BEAMS 
infrastructure to operate. This section details SF-BEAMS in general, and
presents the requirements for data persistence after classifying NetBEAMS based
on the taxonomies defined in the previous chapter.

@@

In general, the SF-BEAMS network infrastructure contains a varying number of
wired and wireless devices attached to pylons at the pier. Each of them
is responsible for observing different conditions of the area, having its own
mechanisms for internal storage for the collected data. In this way, data can
be directly transferred to the labs via Ethernet cables or collected manually
with a laptop computer. The current data collection process for SF-BEAMS is
described in Figure \ref{fig:SF-BEAMS-system-architecture} and documented at
\cite{sfbeams-current-system}. Upon collecting data from sensors, the RTC
staff uses automation scripts written in Matlab \cite{matlab} to process,
index and distribute the raw data in different formats. One of these formats
is the OPeNDAP \cite{opendap}, which is widely used at research institutions
to promote easier data exchange amongst them. The SF-BEAMS website provides
access to the collected data through the Internet at
http://sfbeams.sfsu.edu:8080/opendap. An example of an HTTP Request from
accessing the data for a specific device is shown in Listing
\ref{file:rtc-ysi-opendap} on page \pageref{file:rtc-ysi-opendap}.

@@

As described in Section \ref{sec:sn-infrastructure}, sensor devices produce
the observed data based on properties of measurements defined by its
manufacturer. SF-BEAMS has used, among other devices, the ``6600 ESD V2" Sonde
\cite{YSI-Sonde}, manufactored by YSI Incorporated, as seen in Figure
\ref{fig:ysi-device}. Although it is a small device of under 55 cm in length,
8.9 cm in diameter, and weighting approximately 3.18 kg, it is a powerful water
quality-monitoring device, capable of producing around 52 bytes on a single
real-time data stream reading regarding 12 different parameters such as water
temperature, salinity, turbiny, among others. An example of the data stream is
as shown in Listing \ref{data:ysi-stream} on page \pageref{data:ysi-stream}.

@@

\lstset{label=data:ysi-stream,caption=The Collected Data Stream from the
YSI Sonde 6600ESDV2}
\begin{lstlisting}
12.20    192    179 5588.40   0.09   0.084   0.059  7.98   -79.6   99.5  8.83  0.4     8.7
\end{lstlisting}

In order to have access to the observed data, the RTC network staff downloads
the data using one of the device's connections such as the RS-232 serial connector
\cite{rs232}. Then, they upload the data into their data servers where they are
processed and archived for historical analysis. According to one
staff member, the infrastructure of the SF-BEAMS is comprised of the 
following components (as of May 2009):

@@

Considering the infrastructure, the data load in the server-side
of RTC's laboratories produced by the YSI Sonde sensors devices can be
estimated by analyzing the data distribution over a period of one year, as
shown in Table \ref{tab:year-data-distribution}.

\begin{table}[!b]
    \label{tab:year-data-distribution}
     \begin{center}
      \begin{tabular}{|c|c|c|c|c|c|c|}\hline 
        \textbf{\# of YSIs} & \textbf{Rate} & \textbf{Hourly} &
        \textbf{Daily} & \textbf{Weekly} & \textbf{Monthly} & \textbf{Yearly}\\\hline 
        1 & 1 min & 3.04 Kb & 73.12 Kb & 511.87 Kb & 1.99 Mb & 23.99 Mb\\\hline 
        5 & 6 min & 15.23 Kb & 365.62 Kb & 2.5 Mb & 9.99 Mb & 119.97 Mb\\\hline 
        1 & 15 min & 0.5 Kb & 12.18 Kb & 85.31 Kb & 341.25 Kb & 3.99 Mb\\\hline 
        5 & 1 min & 2.54 Kb & 60.93 Kb & 426.56 Kb & 1.67 Mb & 19.99 Mb\\\hline
        1 & 6 min & 0.2 Kb & 4.87 Kb & 34.12 Kb & 136.5 Kb & 1.6 Mb\\\hline 
        5 & 15 min & 1.0 Kb & 24.37 Kb & 170.62 Kb & 682.5 Kb & 7.99 Mb\\\hline
        \end{tabular}
        \caption{Estimated data volume produced by the RTC's YSI sondes}
      \end{center}
\end{table}

@@

\begin{itemize}
  \item \textbf{The Purpose of Sensor Data}: the data from SF-BEAMS is exclusively
  used for the purpose of Data Archival, since they are stored in RTC's lab;
  \item \textbf{The Location of the Sensor Data}: it is clear that the purpose
  of the data is directly related to the sensors location and, thus, the
  strategy of \textbf{External Data Storage} is used to store its collected
  data at the network sink at the RTC site;
  \item \textbf{Data Model}: since the format used by the RTC staff to share
  data with researchers is in the OPeNDAP format, the data
  model used is the tabular data model, since it uses the comma-delimited
  files. In this way, the \textbf{Schema-less} model is used;
  \item \textbf{Data Description}: once the collected data reaches the RTC
  lab, a quality assurance process is executed, and the data is transformed
  into the format used by OPeNDAP. The process to describe the collected data
  includes the time of the data collection, therefore, one of the \textbf{Time
  Dimensions} such as the valid time is used. Since the files are stored with
  specific file name structure, using timestamps to describe the data, the use
  of transaction time may be also considered. Finally, the contents of the
  files contain the \textbf{Data Identity} as shown in the samples of the
  collected data in Listing \ref{file:rtc-ysi-opendap} on page
  \pageref{file:rtc-ysi-opendap};
  \item \textbf{Query Processing Mechanism}: similar to the data stored in a
  centralized way, the query processing is also \textbf{Centralized}, as the
  data is stored on file system;
  \item \textbf{Data Volume}: the volume of data produced is characterized by small.
 It is related to the volume of data produced by the sensor devices,
  which is numerical data streams, and the total number of sensor devices;
  \item \textbf{System Organization}: SF-BEAMS runs on a Single System, as
  the collected data is stored directly in the file-system after previous 
  processing. However, the collected data is provided to the Internet using
  the OPeNDAP Hyrax system as a middleware.
\end{itemize}

@@

Although the execution of RTC's sensor network can be operated as it
was mentioned in the previous sections, \cite{netbeams2009} described some of the
operational challenges faced by the RTC staff during regular activities of the
data collection process. It was clear to the research group that SF-BEAMS
could have not only its data gathering process automated, but also its data
management and distribution. By the use of COTS\footnote{Common-Off-The-Shelf
designates a product that is produced and sold in bulk} embedded devices and
open-source \cite{open-source} software, the research group developed NetBEAMS,
component-based approach which suggested the improvement of operational
activities from SF-BEAMS using systems automation. In brief, one of the
ongoing problems of NetBEAMS was regarding Data Persistence, the primary
motivation of this report.

@@

The scope of this work is defined as to provide a persistence layer for
NetBEAMS, making a selection of a persistence layer based on the taxonomical
classification of SF-BEAMS. Furthermore, the technology to be selected for the
persistence layer must take into account the requirements of NetBEAMS primary
users such as Computer Science researchers, as well as end users such as
Biologists from the RTC laboratories. The requirements for the data
persistence specifically for NetBEAMS are described in the following section.
